---
title: "Reproducible Figures Assignment: Question 2"
output:
  html_document: default
---

### Introduction

In this assignment, I will develop and test hypotheses using the Palmer Penguins dataset from Horst AM et al., 2020^1^, who studied 344 penguins across three islands of the Palmer Archipelago, Antarctica.

<font size="1"> ^1^Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. doi:10.5281/zenodo.3960218, R package version 0.1.0, https://allisonhorst.github.io/palmerpenguins/. </font>

```{r Data Exploration}
#Load libraries
library(ggplot2)
library(palmerpenguins)
library(janitor)
library(dplyr)
library(ggpubr)

head(penguins_raw) #Look at the first 6 row of the penguins_raw data

#Defining functions:

# A function to clean column names
clean_column_names <- function(penguins_data) {
  penguins_data %>%
    clean_names()
}

# A function to shorten species names
shorten_species <- function(penguins_data) {
  penguins_data %>%
    mutate(species = case_when(
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo"
    ))
}

# A function to remove empty columns or rows
remove_empty_columns_rows <- function(penguins_data) {
  penguins_data %>%
    remove_empty(c("rows", "cols"))
}

#Cleaning the data
penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows()

head(penguins_clean)

```
```{r}
#Create a function for background colours
background.theme <- function() {
   theme(panel.border = element_rect(color= "#3B3A3A", fill = NA), #change color of the panel 
    panel.background = element_rect(fill = "#fff5e5", color = "#fff5e5"),
    plot.background = element_rect(fill = "#fff5e5", color = "#fff5e5")
    
   )
    
  }

#Exploratory Figure
exploratory.figure <- ggplot(data = penguins_clean, 
                             aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point()+
  labs(title = "Figure. 1")+ #plot title
  theme(plot.title = element_text(hjust = 0.5)) + #title in center
  theme(text = element_text(family = "Times New Roman", size = 15, color = "#033ca6"), #change font
        plot.title = element_text(
          family = "Times New Roman", 
          size = 17,
          face = "bold"))+
  background.theme()

exploratory.figure #Exploratory figure plotting body_mass_g against flipper_length_mm
```

Figure. 1 shows an exploratory figure, displaying body mass (g) plotted against flipper length (mm).

### Hypothesis

Null hypothesis: There is no correlation between body mass and flipper length

Alternative hypothesis: There is a correlation between body mass and flipper length

### Statistical Methods

I will use Pearson’s Correlation Coefficient (r) to calculate the linear correlation between body mass and flipper length in penguins.

```{r Statistics}

# Make sure your code prints. 
results.figure <- ggscatter(penguins_clean, x = "body_mass_g", y = "flipper_length_mm", title = "Figure. 2",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson", #calculate Pearson’s Correlation Coefficient
          add.params = list(color = "blue"), #change color of linear regression line
          xlab = "Body Mass (g)", ylab = "Flipper Length (mm)")+
  theme(plot.title = element_text(hjust = 0.5)) + #title in center
  theme(text = element_text(family = "Times New Roman", size = 15, color = "#033ca6"), #change font
        plot.title = element_text(
          family = "Times New Roman", 
          size = 17,
          face = "bold"))+
  background.theme()

```

### Results & Discussion

My results show that there is a significant, positive correlation between body mass and flipper length in penguins from the Palmer Archipelago. These results are displayed in Figure. 2, which shows a scatter plot of body mass and flipper length with the linear regression line and 95% confidence interval in blue. The coefficient value (R = 0.87) is greater than 0.5, showing that this association is a strong correlation and the p value is small (p < 2.2e^-16^) suggesting that it's unlikely that this result was due to chance. 

This result is in line with my expectations as greater body mass is likely needed to support greater flipper lengths, and an increase in flipper length contributes to an increase in body mass. Future analyses could compare the strength of this relationship in different species of penguin.

```{r Plotting Results}

results.figure

```

### Conclusion

To conclude, there is a high degree of positive correlation between body mass and flipper length in penguins from the Palmer Archipelago.

